Refinements for open automata

### 2023-02-24 2021-Refinement/paper/refinementarticleV2.tex
sketch proof

\subsection{A bisimulation for open automata.}

@@

\begin{definition}[Non-blocking composition]
Let \(A_i = \OAg[i]\) with \( 0 \leq i \leq n\) be a family of open automata.
Let \(A = A_0\mdbrk{j_i \mapsto A_i \middle| 1 \leq i \leq n}\) and $A= \OAg$.

@@

This is valid for all $s_{2p} \in (s_{2p})^{p\in P}$, then we have:
\begin{multline*}
 V_1 \uplus V_2 \uplus V_3  \vdash \underbrace{\bigvee_{{p\in P}}\mpar{R_{12}\mpar{s_1, s_{2p}} \wedge R_{23}\mpar{s_{2p},s_3}}}_{R_{13}(s_1,s_3)} \implies  ( \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} )
 \lor¢ \lnot( \bigvee_\subbox{t_3 \in \fOT{s_3}} \fguard{t_3} )
\end{multline*}
By removing the $A_2$ variables that have not effect on $A_2$ et $A_3$,
we get the desired result:
\begin{multline*}
 V_1 \uplus  V_3  \vdash R_{13}(s_1,s_3) \implies  ( \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} )
 \lor¢ \lnot( \bigvee_\subbox{t_3 \in \fOT{s_3}} \fguard{t_3} )
\end{multline*}
\qed
\end{enumerate}
\section{Proof of Context Refinement (Theorem~\ref{thm:ContextRefinement})}
\begin{quote}
Suppose that $\wrel{A_1}{A_2}{H}$, \(k \in H\) \cancel{and that \(A_1\subst{A_3}{k}\) is non-blocking}. We have: 
\[\wrel{A_1\subst{A_3}{k}}{A_2\subst{A_3}{k}}{J_3 \uplus H \setminus \mbrc{k}}\]
\end{quote}
\smallskip

@@

Let's start  with the hypothesis of the theorem stating that $\wrel{A_1}{A_2}{H}$, thus we have for all  $\mpar{s_1, s_2} \in S_1 \times S_2$ 
\begin{multline*}
V_1 \uplus V_2 \vdash R\mpar{s_1, s_2} \nwedge \neg (\bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1}) \implies   \neg(\bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2})
\end{multline*}
Because the guards from transitions of $A_1$ are still transitions of $A_{13}$, we get:  
\begin{multline*}
V_{13}   \vdash \neg(\bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1}) \implies \neg(\bigvee_\subbox{t_{13} \in \fOT{s_{13}}} \fguard{t_{13}})
\end{multline*}

Similarly, transitions of $A_2$ are still transitions of $A_{23}$,  we get:
\begin{multline*}
V_{23}   \vdash \neg(\bigvee_\subbox{t_1 \in \fOT{s_2}} \fguard{t_2}) \implies \neg(\bigvee_\subbox{t_{23} \in \fOT{s_{23}}} \fguard{t_{23}})
\end{multline*}
From the three previous formulas we deduce:
  \begin{multline*}
V_{13} \uplus V_{23} \vdash R\mpar{s_1, s_2} \nwedge \neg (\bigvee_\subbox{t_{13} \in \fOT{s_{13}}} \fguard{t_{13}}) \implies   \neg(\bigvee_\subbox{t_{23} \in \fOT{s_{23}}} \fguard{t_{23}})
\end{multline*}

which also implies:
\begin{multline*}
V_{13} \uplus V_{23}  \vdash \\ \underbrace{R\mpar{s_1, s_2} \wedge  \mpar{\,\,\bigwedge_\subbox{v_3\in V_3}  v^1_3=v^2_3 \wedge s^1_3=s^2_3}}_{R^{\prime}(s_{13},s_{23})}  \wedge \neg(\bigvee_\subbox{t_{13} \in \fOT{s_{13}}} \fguard{t_{13}})   \implies \neg(\bigvee_\subbox{t_{23} \in \fOT{s_{23}}} \fguard{t_{23}})
\end{multline*}
which gives us what we needed to demonstrate. \qed

\end{enumerate}

@@

Let $R$ be the refinement relation relating states of $A_2$ and $A_3$. 
Let us denote with $t^2$ and $t^3$  the elements of $A_2$ and $A_3$ respectively.
Consider any two states $s_{12} = \mpar{s_1,s^1_2}$ and $s_{13} = \mpar{s^2_1,s_3}$ ($s^1_1$ and $s^2_1$  are the same with renaming). We define a relation $R'$ relating states of $s_{12}$ and $s_{13}$ as follows:
\[ R' \mpar{s_{12}, s_{13}} = R\mpar{s_{2}, s_{2}} \wedge \bigwedge_\subbox{v_1\in V_1}
 v^1_1=v^2_1\wedge s^1_1 = s^2_1\]
Let us denote $H'=H\cup J_1 \setminus\{k\}$.
We want to prove that $\mpar{R', H'}$  is a hole-tracking simulation of $A_{12}$ and $A_{13}$. As for the previous proof we deal with cases:
\begin{enumerate}
\item \TODO{Pareil que le theorem precedent}

@@

\begin{align*}
		&  \OTx{12}{}{12}{12} \in \fOT{s_{12}}
\end{align*}		
there exists an indexed family $t_{13x}$ of OTs originating from $s_{13}$ that simulates it: 
\begin{align*}
		&  \mpar{\OTx{13}{x}{13x}{13x} \in \fOT{s_{13}}}^{x \in X}
\end{align*}	
such that  $\mpar{\forall x \in X, J'_{13x} \cap H' = J'_{12} \cap H'}$ and
\begin{multline*}	
V_{12} \uplus V_{13} \uplus \fvars{t_{12}} \vdash\\ R'\mpar{s_{12}, s_{13}} \wedge g_{12} \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\everymath{\displaystyle}\begin{array}{l}
			\alpha_{12} = \alpha_{13x} \wedge \bigwedge_\subbox{j \in J'_{13x} \cap H'} \beta_{12j} = \beta_{13xj} \wedge\\[12pt]
			 g_{13x} \wedge R'\mpar{s'_{12}, s'_{13x}}\psubst{\psi_{12} \uplus \psi_{13x}}
		\end{array}} 
	\end{multline*}
By definition of composition and open automata refinement we have: 
\begin{align*}
&V_{12}= V^1_1\uplus V_3 \text{ and }
V_{13}= V^2_1\uplus V_3\\
& H'=H_{23}\uplus J_1 \setminus \{k\} \subseteq J_1\uplus \mpar{J_2 \cap J_3} \setminus \{k\} = \mpar{J^1_1\uplus J_2 \setminus \{k\} } \cap \mpar{J^2_1\uplus J_3 \setminus \{k\}} = J_{12 } \cap J_{13} 
\end{align*}
We have by hypothesis $\wrel{A_2}{A_3}{H}$, then for any open transition $t_2$ in $T_2$  originating from $s_2$:
\begin{align*}
\OTx{2}{}{2}{2} \in \fOT{s_2}
\end{align*}
there exists an indexed family of OTs originating from $s_{2}$: 
\begin{align*}
\mpar{\OTx{3}{x}{3x}{3x} \in \fOT{s_{3}}}^{x \in X} 
\end{align*}
such that $\forall x \in X, J'_{3x} \cap H = J'_2 \cap H$ and
\begin{multline*}
%\everymath{\displaystyle}
V_2 \uplus V_3 \uplus \fvars{t_2} \vdash\\ R\mpar{s_2, s_3} \wedge g_2 \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\everymath{\displaystyle}
\begin{array}{l}
			\alpha_2 = \alpha_{3x} \wedge \bigwedge_\subbox{j \in J'_{3x} \cap H} \beta_{2j} = \beta_{3xj} \nwedge\\
			 g_{3x} \wedge R\mpar{s'_2, s'_{3x}}\psubst{\psi_2 \uplus \psi_{3x}}
		\end{array}}  \qquad \quad (*)
\end{multline*}	
\smallskip

@@

\item Lastly, we have to prove the satisfaction of the deadlock reduction condition, i.e., for all $\mpar{s_{12}, s_{13}} \in S_{12} \times S_{13}$
\begin{multline*}
V_{12} \uplus V_{13}  \vdash \\ R\mpar{s_{12}, s_{13}} \nwedge  \neg(\bigvee_\subbox{t_{12} \in \fOT{s_{12}}} \fguard{t_{12}}) \implies   \neg (\bigvee_\subbox{t_{13} \in \fOT{s_{13}}} \fguard{t_{13}}) 
\end{multline*}

Let's start  with the hypothesis of the theorem stating that $\wrel{A_2}{A_3}{H}$, thus we have for all  $\mpar{s_2, s_3} \in S_2 \times S_3$ 
\begin{multline*}
V_2 \uplus V_3  \vdash R\mpar{s_2, s_3} \nwedge \neg (\bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2}) \implies   \neg(\bigvee_\subbox{t_3 \in \fOT{s_3}} \fguard{t_3})
\end{multline*}
This implies: 
\begin{multline*}
V_2 \uplus V_3  \vdash R\mpar{s_2, s_3} \nwedge \neg (\bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2}) \vee \neg (\bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1}) \implies  \\ \neg(\bigvee_\subbox{t_3 \in \fOT{s_3}} \fguard{t_3}) \vee \neg (\bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1})
\end{multline*}

\RAB{Trop trivial...}
~\qed
\end{enumerate}

>>>>

\subsection{A bisimulation for open automata}

@@

\begin{definition}[Non-blocking composition] \label{Def:Non-block}
Let \(A_i = \OAg[i]\) with \( 0 \leq i \leq n\) be a family of open automata.
Let \(A = A_0\mdbrk{j_i \mapsto A_i \middle| 1 \leq i \leq n}\) and $A= \OAg$.

@@

This is valid for all $s_{2p} \in (s_{2p})^{p\in P}$, then we have:
\begin{multline*}
 V_1 \uplus V_2 \uplus V_3  \vdash \underbrace{\bigvee_{{p\in P}}\mpar{R_{12}\mpar{s_1, s_{2p}} \wedge R_{23}\mpar{s_{2p},s_3}}}_{R_{13}(s_1,s_3)} \implies  ( \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} )
 \lor¢ \lnot( \bigvee_\subbox{t_3 \in \fOT{s_3}} \fguard{t_3} )
\end{multline*}
By removing the $A_2$ variables that have not effect on $A_2$ et $A_3$,
we get the desired result:
\begin{multline*}
 V_1 \uplus  V_3  \vdash R_{13}(s_1,s_3) \implies  ( \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} )
 \lor¢ \lnot( \bigvee_\subbox{t_3 \in \fOT{s_3}} \fguard{t_3} )
\end{multline*}
\qed
\end{enumerate}
\section{Proof of Context Refinement (Theorem~\ref{thm:ContextRefinement})}
\begin{quote}
Suppose that $\wrel{A_1}{A_2}{H}$, \(k \in H\) and that \(A_1\subst{A_3}{k}\) is non-blocking. We have: 
\[\wrel{A_1\subst{A_3}{k}}{A_2\subst{A_3}{k}}{J_3 \uplus H \setminus \mbrc{k}}\]
\end{quote}
\smallskip

@@

By hypothesis  the second  \(A_1\subst{A_3}{k}\) is non-blocking,  then according to the Definition \ref{Def:Non-block} we have for all state $s_{13}$ which is reachable, i.e. $ \fvars{t_{13}}\vdash \reach{A_{13}}\mpar{s_{13}}$,  the following:

\begin{multline*}
V_{13} \uplus \biguplus_\subbox{t_1 \in \fOT{s_1}} \fvars{t_1}  \vdash \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} \implies \bigvee_\subbox{t_{13} \in \fOT{s_{13}}} \fguard{t_{13}}
\end{multline*}
From this we can infer:
\begin{multline*}
V_{13}  \uplus \biguplus_\subbox{t_1 \in \fOT{s_1}} \fvars{t_1} \vdash  \neg(\bigvee_\subbox{t_{13} \in \fOT{s_{13}}} \fguard{t_{13}})  \implies   \neg(\bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1})
\end{multline*}

Furthermore, from the second  hypothesis  stating that $\wrel{A_1}{A_2}{H}$,  we have for all  $\mpar{s_1, s_2} \in S_1 \times S_2$ 
\begin{multline*}
V_1 \uplus V_2 \vdash R\mpar{s_1, s_2} \nwedge \neg (\bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1}) \implies   \neg(\bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2})
\end{multline*}

From the two previous formula can deduce:
\begin{multline*}
V_{13} \uplus V_2 \uplus \biguplus_\subbox{t_1 \in \fOT{s_1}} \fvars{t_1} \vdash\\ R\mpar{s_1, s_2} \nwedge \neg(\bigvee_\subbox{t_{13} \in \fOT{s_{13}}} \fguard{t_{13}})  \implies   \neg(\bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2})
\end{multline*}

Because the guards from transitions of $A_2$ are still transitions of $A_{23}$,  thus we have:
\begin{multline*}
V_{23}   \vdash \neg(\bigvee_\subbox{t_1 \in \fOT{s_2}} \fguard{t_2}) \implies \neg(\bigvee_\subbox{t_{23} \in \fOT{s_{23}}} \fguard{t_{23}})
\end{multline*}

From the two previous formula and by introducing the following tautology\\ $ \mpar{\,\,\displaystyle \bigwedge_\subbox{v_3\in V_3} v^1_3=v^2_3\wedge s^1_3 = s^2_3}$, we get:
\begin{multline*}
V_{13} \uplus V_{23}  \vdash  \underbrace{R\mpar{s_1, s_2} \wedge  \mpar{\,\,\bigwedge_\subbox{v_3\in V_3}  v^1_3=v^2_3 \wedge s^1_3=s^2_3}}_{R^{\prime}(s_{13},s_{23})} \wedge \neg(\bigvee_\subbox{t_{13} \in \fOT{s_{13}}} \fguard{t_{13}})   \\ \implies \neg(\bigvee_\subbox{t_{23} \in \fOT{s_{23}}} \fguard{t_{23}})
\end{multline*}
which gives the expected result. \qed
\end{enumerate}

@@

Let $R$ be the refinement relation relating states of $A_2$ and $A_3$. 
Let us denote with $t^2$ and $t^3$  the elements of $A_2$ and $A_3$ respectively.
Consider any two states $s_{12} = \mpar{s_1,s^1_2}$ and $s_{13} = \mpar{s^2_1,s_3}$ ($s^1_1$ and $s^2_1$  are the same with renaming). We define a relation $R'$ relating states of $s_{12}$ and $s_{13}$ as follows:
\[ R' \mpar{s_{12}, s_{13}} = R\mpar{s_{2}, s_{2}} \wedge \bigwedge_\subbox{v_1\in V_1}
 v^1_1=v^2_1\wedge s^1_1 = s^2_1\]
Let us denote $H'=H\cup J_1 \setminus\{k\}$.
We want to prove that $\mpar{R', H'}$  is a hole-tracking simulation of $A_{12}$ and $A_{13}$. As for the previous proof we deal with cases:
\begin{enumerate}
\item \TODO{I skip it, I think the same as previous theorem}

@@

\begin{align*}
		&  \OTx{12}{}{12}{12} \in \fOT{s_{12}}
\end{align*}		
there exists an indexed family $t_{13x}$ of OTs originating from $s_{13}$ that simulates it: 
\begin{align*}
		&  \mpar{\OTx{13}{x}{13x}{13x} \in \fOT{s_{13}}}^{x \in X}
\end{align*}	
such that  $\mpar{\forall x \in X, J'_{13x} \cap H' = J'_{12} \cap H'}$ and
\begin{multline*}	
V_{12} \uplus V_{13} \uplus \fvars{t_{12}} \vdash\\ R'\mpar{s_{12}, s_{13}} \wedge g_{12} \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\everymath{\displaystyle}\begin{array}{l}
			\alpha_{12} = \alpha_{13x} \wedge \bigwedge_\subbox{j \in J'_{13x} \cap H'} \beta_{12j} = \beta_{13xj} \wedge\\[12pt]
			 g_{13x} \wedge R'\mpar{s'_{12}, s'_{13x}}\psubst{\psi_{12} \uplus \psi_{13x}}
		\end{array}} 
	\end{multline*}
By definition of composition and open automata refinement we have: 
\begin{align*}
&V_{12}= V^1_1\uplus V_3 \text{ and }
V_{13}= V^2_1\uplus V_3\\
& H'=H_{23}\uplus J_1 \setminus \{k\} \subseteq J_1\uplus \mpar{J_2 \cap J_3} \setminus \{k\}= \mpar{J^1_1\uplus J_2 \setminus \{k\} } \cap \mpar{J^2_1\uplus J_3 \setminus \{k\}} \\ &\,\,\,\, = J_{12 } \cap J_{13} 
\end{align*}
We have by hypothesis $\wrel{A_2}{A_3}{H}$, then for any open transition $t_2$ in $T_2$  originating from $s_2$:
\begin{align*}
\OTx{2}{}{2}{2} \in \fOT{s_2}
\end{align*}
there exists an indexed family of OTs originating from $s_{2}$: 
\begin{align*}
\mpar{\OTx{3}{x}{3x}{3x} \in \fOT{s_{3}}}^{x \in X} 
\end{align*}
such that $\forall x \in X, J'_{3x} \cap H = J'_2 \cap H$ and
\begin{multline*}
%\everymath{\displaystyle}
V_2 \uplus V_3 \uplus \fvars{t_2} \vdash\\ R\mpar{s_2, s_3} \wedge g_2 \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\everymath{\displaystyle}
\begin{array}{l}
			\alpha_2 = \alpha_{3x} \wedge \bigwedge_\subbox{j \in J'_{3x} \cap H} \beta_{2j} = \beta_{3xj} \nwedge\\
			 g_{3x} \wedge R\mpar{s'_2, s'_{3x}}\psubst{\psi_2 \uplus \psi_{3x}}
		\end{array}}  \qquad \quad (*)
\end{multline*}	
\smallskip

@@

\item Lastly, we have to prove the satisfaction of the deadlock reduction condition, i.e., for all $\mpar{s_{12}, s_{13}} \in S_{12} \times S_{13}$
\begin{multline*}
V_{12} \uplus V_{13}  \vdash  R\mpar{s_{12}, s_{13}} \nwedge  \neg(\bigvee_\subbox{t_{12} \in \fOT{s_{12}}} \fguard{t_{12}}) \implies   \neg (\bigvee_\subbox{t_{13} \in \fOT{s_{13}}} \fguard{t_{13}}) 
\end{multline*}

Let's start  with the hypothesis of the theorem stating that $\wrel{A_2}{A_3}{H}$, thus we have for all  $\mpar{s_2, s_3} \in S_2 \times S_3$ 
\begin{multline*}
V_2 \uplus V_3  \vdash R\mpar{s_2, s_3} \nwedge \neg (\bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2}) \implies   \neg(\bigvee_\subbox{t_3 \in \fOT{s_3}} \fguard{t_3})
\end{multline*}

By adding the same terms to both sides of the implication, we get:
\begin{multline*}
V_2 \uplus V_3  \vdash R\mpar{s_2, s_3} \nwedge \neg (\bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2}) \vee \neg (\bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1}) \vee \neg \bigcirc \\ \vspace{4em} \implies   \neg(\bigvee_\subbox{t_3 \in \fOT{s_3}} \fguard{t_3})\vee \neg (\bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1})\vee \neg \bigcirc
\end{multline*}

\RAB{On aurait pas envie d'ecrire  l'egalite des actions de trous}

~\qed
\end{enumerate}
